---
title: "Does college major contribute to difference in the income?"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
The libarary containing the dataset is loaded in R environment and the dataset is imported in the memory.  
```{r}
library(dplyr)
library(ggplot2)
library(collegeIncome)
data(college)
```


## Dataset codebook
* rank: Rank by median earnings
* major_code: Major code
* major: Major description
* major_category: Category of major
* total: Total number of people with major
* sample_size: Sample size of full-time, year-round individuals used for income/earnings estimates: p25th, median, p75th
* p25th: 25th percentile of earnings
* median: Median earnings of full-time, year-round workers
* p75th: 75th percentile of earnings
* perc_men: % men with major (out of total)
perc_women: % women with major (out of total)
* perc_employed: % employed (out of total)
* perc_employed_fulltime: % employed 35 hours or more (out of employed)
* perc_employed_parttime: % employed less than 35 hours (out of employed)
* perc_employed_fulltime_yearround: % employed at least 50 weeks and at least 35 hours (out of employed and full-time)
* perc_unemployed: % unemployed (out of employed)
* perc_college_jobs: % with job requiring a college degree (out of employed)
* perc_non_college_jobs: % with job not requiring a college degree (out of employed)
* perc_low_wage_jobs: % in low-wage service jobs (out of total)

## Quick overview of the dataset
The dataset is quickly reviewed using the *head* an *str* function.   
```{r}
head(college)
str(college)
```

A basic analysis of the average income for each category is performed.
```{r}
college_tmp <- college %>% select(major_category, median, p25th, p75th)
college2 <- college_tmp %>% group_by(major_category) %>% summarize(median_mean = mean(median, na.rm = TRUE), p25th_mean = mean(p25th, na.rm = TRUE), p75th_mean = mean(p75th, na.rm = TRUE))
college2 <- college2 %>% arrange(desc(median_mean))
college2
```
The 3 higher median incomes is associated with a major in 1. Business, 2. Biology & Life Sciences and 3. Agriculture & Natural Resources. The 3 lower incomes are Humanities & Liberal Arts, Computers & Mathematics and Interdisciplinary.   


In order to check the distribution for each major category and the p25th, median and p75th income, a boxplot is generated. The distribution of each quantiles per major category do not show any trend, suggesting that the income is independent of the major. 
```{r}
major_cat <- rep(college_tmp$major_category, times = 3)
income_cat <- rep(c("median", "p25th", "p74th"), each = length(college_tmp$major_category))
income <- c(college_tmp$median, college_tmp$p25th, college_tmp$p75th)
college_tmp2 <- data.frame(cbind(major_cat, income_cat, income))
college_tmp2[,3] <- as.numeric(college_tmp2[,3])
g <- ggplot(college_tmp2, aes(income, major_cat))
g + geom_boxplot() + facet_grid(income_cat~.)
```


## Regression model
A regression models is generateed from the dataset where the median as the outcome and the major category as the factor. 

```{r}
fit <- lm(median ~ major_category, college_tmp)
round(summary(fit)$coef, 3)
plot(fit)
```


The intercept is the major category Agriculture & Natural Resources. All the different variables are compared to the intercept, the p-value is not statistically significant compared to each category. The plot of the residuals doesn't show any pattern and looks normally distributed. Let's use a multiple regression model to confirm this observation.  
```{r}
anova(fit)
```


When comparing all the college majors against each other, the anova fail to find any significant difference in the mean income.

## Special case
Let's compare now the median income from the major category business compared to the rest. 

```{r}
major_cat2 <- relevel(as.factor(college_tmp$major_category), "Business" )
fit2 <- lm(median ~ major_cat2, college_tmp)
summary(fit2)
par(mfrow = c(2,2))
plot(fit2)
```


By comparing the major category **Business** to the other categories, we can observe a clear difference with some major categories. There is a significant difference (p-value < 0.05) when compared to **Arts**, **Computers and Mathematics**, **Education**, **Engineering**, **Humanities & Liberal Arts** and **Social Science**.  



## Conclusion
In order to answer the question below, a basic analysis was performed.   

**Question:** Based on your analysis, would you conclude that there is a significant association between college major category and income?  


The assumptions could have been stated as the following:
* NULL hypothesis: there is no difference in the mean of the incomes when holding a specific major.
* Alternative hypothesis: There is a difference in the mean of the incomes between major categories.


Our basic analysis shows that there is no difference in the income when holding a specific major. However, when orienting the analysis to the higher income a difference between some majors can be observed. There is a significant difference (p-value < 0.05) btween **Business** compared to **Arts**, **Computers and Mathematics**, **Education**, **Engineering**, **Humanities & Liberal Arts** and **Social Science**. These results need to be adjusted in order to confirm such trend. 

